KiCS2: A New Compiler from Curry to Haskell

### 2011-04-05 papers/wflp2011/paper.tex

\begin{figure}
\centering
\begin{tabular}{|@{~~}l@{~~}|*{4}{@{~~}r@{~~}|}}
\hline
System  & ReverseUser & Reverse &  Tak  & TakPeano \\\hline
KiCS2   &        0.13 &    0.11 &  0.24 &     0.82 \\
PAKCS   &        2.07 &    1.91 & 39.19 &    60.69 \\
MCC     &        0.37 &    0.38 &  1.03 &     3.99 \\
GHC     &        0.11 &    0.10 &  0.04 &     0.50 \\
SICStus &        0.39 &    0.28 &  0.50 &     4.82 \\
SWI     &        1.33 &    1.14 &  1.87 &    12.72 \\
\hline
\end{tabular}
\caption{Benchmarks: first-order functional programs}
 \label{fig:bench-first-order}
\end{figure}
%
The first collection of benchmarks (Fig.~\ref{fig:bench-first-order})
are purely first-order functional programs.
The Prolog (SICStus, SWI) and Haskell (GHC) programs have been rewritten
according to the Curry formulation.
``ReverseUser'' is the naive reverse program applied to a list of 4096 elements,
where all data (lists, numbers) are user-defined.
``Reverse'' is the same but with built-in lists.
``Tak'' is a highly recursive function on naturals applied
to arguments (27,16,8) and ``TakPeano'' is the same but with
user-defined natural numbers in Peano representation.
Note that the Prolog programs use a strict evaluation strategy
in contrast to all others. Thus, the difference between PAKCS and SICStus
shows the overhead to implement lazy evaluation in Prolog.

@@

\begin{figure}
\centering
\begin{tabular}{|@{~~}l@{~~}|*{3}{@{~~}r@{~~}|}}
\hline
Method               & PermSort & PermSortPeano &  NDNums \\\hline
\code{printValsDFS}  &    2.61  &         3.99  & $\infty$ \\
depth-first search   &    5.10  &         7.62  & $\infty$ \\
breadth-first search &   18.54  &        20.23  &    10.96 \\
iterative deepening  &    8.33  &        10.01  &     3.10 \\
%parallel (1 proc.)   &    5.84  &         6.54  & $\infty$ \\
%parallel (2 proc.)   &    3.46  &         3.79  & $\infty$ \\
\hline
\end{tabular}
\caption{Benchmarks: comparing different search strategies}
 \label{fig:search-strategies}
\end{figure}
%
In order to compare the various search strategies,
Fig.~\ref{fig:search-strategies} contains some corresponding
benchmarks.
``PermSort'' and ``PermSortPeano'' are the programs
discussed in Fig.~\ref{fig:bench-nondet} and ``NDNums'' is the program
\begin{curry}
  f n = f (n+1) ? n
\end{curry}
where we look for the first solution of \ccode{f 0 == 25000}
(obviously, depth-first search strategies do not terminate on this
expression).
All strategies except for the ``direct print'' method \code{printValsDFS}
translate choice structures into monadic list structures
in order to print them according to the user options.
The benchmarks show that the overhead of this transformation
is acceptable so that this more flexible approach is the default one.

>>>>

\begin{figure}
\centering
\begin{tabular}{|@{~~}l@{~~}|*{4}{@{~~}r@{~~}|}}
\hline
System  & ReverseUser & Reverse &  Tak  & TakPeano \\\hline
KiCS2   &        0.13 &    0.11 &  0.24 &     0.82 \\
PAKCS   &        2.07 &    1.91 & 39.19 &    60.69 \\
MCC     &        0.37 &    0.38 &  1.03 &     3.99 \\
GHC     &        0.11 &    0.10 &  0.04 &     0.50 \\
SICStus &        0.39 &    0.28 &  0.50 &     4.82 \\
SWI     &        1.33 &    1.14 &  1.87 &    12.72 \\
\hline
\end{tabular}
\caption{Benchmarks: first-order functional programs}
 \label{fig:bench-first-order}
\end{figure}
%
The first collection of benchmarks\footnote{All benchmarks are available
at \url{http://www.informatik.uni-kiel.de/~mh/repos/kics2/benchmarks.tgz}.}
(Fig.~\ref{fig:bench-first-order})
are purely first-order functional programs.
The Prolog (SICStus, SWI) and Haskell (GHC) programs have been rewritten
according to the Curry formulation.
``ReverseUser'' is the naive reverse program applied to a list of 4096 elements,
where all data (lists, numbers) are user-defined.
``Reverse'' is the same but with built-in lists.
``Tak'' is a highly recursive function on naturals applied
to arguments (27,16,8) and ``TakPeano'' is the same but with
user-defined natural numbers in Peano representation.
Note that the Prolog programs use a strict evaluation strategy
in contrast to all others. Thus, the difference between PAKCS and SICStus
shows the overhead to implement lazy evaluation in Prolog.

@@

\begin{figure}
\centering
\begin{tabular}{|@{~~}l@{~~}|*{3}{@{~~}r@{~~}|}}
\hline
Method               & PermSort & PermSortPeano &  NDNums \\\hline
\code{printValsDFS}  &    2.61  &         3.99  & $\infty$ \\
depth-first search   &    4.78  &         5.90  & $\infty$ \\
breadth-first search &   18.35  &        20.23  &    10.96 \\
iterative deepening  &    8.33  &        10.01  &     3.10 \\
%parallel (1 proc.)   &    5.84  &         6.54  & $\infty$ \\
%parallel (2 proc.)   &    3.46  &         3.79  & $\infty$ \\
\hline
\end{tabular}
\caption{Benchmarks: comparing different search strategies}
 \label{fig:search-strategies}
\end{figure}
%
In order to compare the various search strategies,
Fig.~\ref{fig:search-strategies} contains some corresponding
benchmarks.
``PermSort'' and ``PermSortPeano'' are the programs
discussed in Fig.~\ref{fig:bench-nondet} and ``NDNums'' is the program
\begin{curry}
  f n = f (n+1) ? n
\end{curry}
where we look for the first solution of \ccode{f 0 == 25000}
(obviously, depth-first search strategies do not terminate on this
expression).
All strategies except for the ``direct print'' method \code{printValsDFS}
translate choice structures into monadic list structures
in order to print them according to the user options.
The benchmarks show that the overhead of this transformation
is acceptable so that this more flexible approach is the default one.
